---
title: "communicating local climate trends"
author: "Victoria"
date: "2023-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries and Functions

```{r}
library(tidyverse)
library(ggimage)
library(here)
library(caladaptr)
library(ggplot2)
library(leaflet)
library(sf)
library(httr)
library(extrafont)

functions_path <- here("..", "climate-scenarios", "build_functions")
source(file.path(functions_path, "get_all_grid_cell_data.R"))
source(file.path(functions_path, "caladapt_map.R"))
source(file.path(functions_path, "find_grid_cell_centroid.R"))
```

## Read in and Wrangle Data for CanESM2 both RCPs

```{r}
# Model CanESM2 RCP 85
ui_grid_cells <- c(46779)
ui_rcp <- 85
ui_gcm <- 'CanESM2'

find_grid_cell_centroid() # creates the dataframe with the above user inputs that the get_all_grid_cell_data() function will use to grab data for
centroids_of_interest %>% pmap(get_all_grid_cell_data, ui_gcm = ui_gcm, ui_rcp = ui_rcp) # grabs data for all grid cells 

CanESM2_85_46779 <- grid_cell_46779 %>% 
  mutate(model_rcp = 'CanESM2_85')

# Model CanESM2 RCP 45
ui_grid_cells <- c(46779)
ui_rcp <- 45
ui_gcm <- 'CanESM2'

find_grid_cell_centroid() # creates the dataframe with the above user inputs that the get_all_grid_cell_data() function will use to grab data for
centroids_of_interest %>% pmap(get_all_grid_cell_data, ui_gcm = ui_gcm, ui_rcp = ui_rcp) # grabs data for all grid cells

CanESM2_45_46779 <- grid_cell_46779 %>% 
  mutate(model_rcp = 'CanESM2_45')
```

## Create Map Graphic

```{r}

```

## Maximum Temperature: Combine Data and Plot

```{r}
daily_data <- CanESM2_85_46779 %>% 
  select(time, max_temp) %>% 
  mutate(max_temp = max_temp * 1.8 + 32)

# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

max_temp_plot <- ggplot() +
  
  # max temp daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = max_temp), color = "#EAE7E5") +
  
  # annotation for daily range
      # label
  annotate("text", x = as.Date("2061-01-01"), y = 60,
           label = "Daily Variation", vjust = -1, color = '#606060', size = 5) +
      # line
  annotate("segment", 
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"), 
           y = 51, yend = 109,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +

  # regression line
    geom_smooth(data = daily_data, aes(x = time, y = max_temp), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # annotation for regression line
  annotate("text", x = as.Date("1969-01-01"), y = 70, 
           label = "Average Trend", vjust = -1, color = '#CC4C34', size = 5) +

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Daily Maximum Temperature (°F) Trends Upward",
    caption = "Data Source: https://cal-adapt.org/data/download/ \n *p-value: < 2.2e-16",
    subtitle = "Daily Maximum Temperature Is Projected to Increase by 0.74°F* each Decade, on Average",
    y = "\nMaximum Temperature (°F)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )  +
  
  # set breaks to match between min temp and max temp plots
  scale_y_continuous(breaks = seq(30, 120, by = 10),
                     sec.axis = dup_axis()
                     )

# export maximum temperature plot
ggsave(filename = "maximum_temperature_plot.png", plot = max_temp_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(max_temp ~ time, data = daily_data)
summary_model <- summary(lm_model)
summary_model
```

## Minimum Temperature: Combine Data and Plot

```{r}
daily_data <- CanESM2_85_46779 %>% 
  select(time, min_temp) %>% 
  mutate(min_temp = min_temp * 1.8 + 32)

# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

min_temp_plot <- ggplot() +
  
  # max temp daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = min_temp), color = "#EAE7E5") +
  
    # annotation for daily range
      # label
  annotate("text", x = as.Date("2061-01-01"), y = 60,
           label = "Daily Variation", vjust = -1, color = '#606060', size = 5) +
      # line
  annotate("segment", 
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"), 
           y = 28, yend = 71,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +

  # regression line
    geom_smooth(data = daily_data, aes(x = time, y = min_temp), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # labels for lines
  #annotate("text", x = unique(trend_data$water_year)[20], y = 60, 
          # label = "Potential Daily Variation", vjust = -1, color = 'white', size = 5) +
  annotate("text", x = as.Date("1969-01-01"), y = 42, 
           label = "Average Trend", vjust = -1, color = '#CC4C34', size = 5) +

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Daily Minimum Temperature (°F) Trends Upward",
    caption = "Data Source: https://cal-adapt.org/data/download/ \n *p-value: < 2.2e-16",
    subtitle = "Daily Minimum Temperature Is Projected to Increase by 0.73°F* each Decade, on Average",
    y = "\nMinimum Temperature (°F)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )  +
  
  # set breaks to match between min temp and max temp plots and add a dual axis
  scale_y_continuous(breaks = seq(10, 80, by = 10),
                     sec.axis = dup_axis()
                     )

# export maximum temperature plot
ggsave(filename = "minimum_temperature_plot.png", plot = min_temp_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(min_temp ~ time, data = daily_data)
summary_model <- summary(lm_model)
summary_model
```

## Maximum Relative Humidity: Combine Data and Plot

```{r}
daily_data <- CanESM2_85_46779 %>%
  select(time, max_humidity) %>% 
  # change units to percents
  mutate(max_humidity = max_humidity * 100)

# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

max_rh_plot <- ggplot() +
  
  # max rh daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = max_humidity), color = "#EAE7E5") +
  
  # annotation for daily range
      # label
  annotate("text", x = as.Date("2061-01-01"), y = 60,
           label = "Daily Variation", vjust = -1, color = '#606060', size = 5) +
      # line
  annotate("segment", 
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"), 
           y = 33, yend = 100,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +
  
    # regression line
  geom_smooth(data = daily_data, aes(x = time, y = max_humidity), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # labels for trend line
  annotate("text", x = as.Date("1969-01-01"), y = 80, 
           label = "Average Trend", vjust = -1, color = '#CC4C34', size = 5) +

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Daily Maximum Relative Humidity (%) Trends Upward",
    caption = "Data Source: https://cal-adapt.org/data/download/ \n *p-value: < 2.2e-16",
    subtitle = "Daily Maximum Relative Humidity Is Projected to Increase by 0.16%* each Decade, on Average",
    y = "\nMaximum Relative Humidity (%)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) + 
  
  # set breaks to match between min rh and max rh plots and add a dual axis
  scale_y_continuous(breaks = seq(0, 100, by = 25),
                     sec.axis = dup_axis()
                     )

# you can see that max humidity has less area in the lower area than in the future. 

# export
ggsave(filename = "maximum_relativehumidity_plot.png", plot = max_rh_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(max_humidity ~ time, data = daily_data)
summary_model <- summary(lm_model)
summary_model
```

## Minimum Relative Humidity: Combine Data and Plot

```{r}
daily_data <- CanESM2_85_46779 %>%
  select(time, min_humidity) %>% 
  # change units to percents
  mutate(min_humidity = min_humidity * 100)

# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

min_rh_plot <- ggplot() +
  
  # max rh daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = min_humidity), color = "#EAE7E5") +
  
  # annotation for daily range
      # label
  annotate("text", x = as.Date("2061-01-01"), y = 60,
           label = "Daily Variation", vjust = -1, color = '#606060', size = 5) +
      # line
  annotate("segment", 
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"), 
           y = 12, yend = 72,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +
  
  # trend line 
  #geom_line(data = trend_data, aes(x = time, y = avg_min_rh), color = '#CC4C34', size = 1) +

  # regression line
  geom_smooth(data = daily_data, aes(x = time, y = min_humidity), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # labels for trend line
  annotate("text", x = as.Date("1969-01-01"), y = 34, 
           label = "Average Trend", color = '#CC4C34', size = 5) +

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Daily Minimum Relative Humidity (%) Trends Upward",
    caption = "Data Source: https://cal-adapt.org/data/download/ \n *p-value: < 2.2e-16",
    subtitle = "Daily Minimum Relative Humidity Is Projected to Increase by 0.17%* each Decade, on Average",
    y = "\nMinimum Relative Humidity (%)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) + 
  
  # set breaks to match between min rh and max rh plots and add a dual axis
  scale_y_continuous(breaks = seq(0, 100, by = 25),
                     sec.axis = dup_axis()
                     )

# export
ggsave(filename = "minimum_relativehumidity_plot.png", plot = min_rh_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(min_humidity ~ time, data = daily_data)
summary_model <- summary(lm_model)
summary_model
```

## Wind: Combine Data and Plot

```{r}

# just CanESM
daily_data <- CanESM2_85_46779 %>% 
  # convert m/s to mph since this is a better known unit
  mutate(wind = wind * 2.23694) %>% 
  select(time, wind)

# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

wind_plot <- ggplot() +
  
  # max rh daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = wind), color = "#EAE7E5") +
  
   # annotation for daily range
      # label
  annotate("text", x = as.Date("2061-01-01"), y = 5,
  label = "Daily Variation", color = '#606060', size = 5) +
      # line
  annotate("segment",
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"),
           y = 2, yend = 27,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +

    # regression line
  geom_smooth(data = daily_data, aes(x = time, y = wind), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # labels for trend line
  annotate("text", x = as.Date("1969-01-01"), y = 10, 
           label = "Average Trend", vjust = -1, color = '#CC4C34', size = 5) +

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Daily Wind Speed (mph) Trends Downward",
    caption = "Data Source: https://cal-adapt.org/data/download/ \n *p-value: < 2.2e-16",
    subtitle = "Daily Wind Speeds Are Projected to Decrease by 0.09 mph* each Decade, on Average",
    y = "\nWind Speed (mph)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  ) + 
  
  # add a dual axis
  scale_y_continuous(sec.axis = dup_axis())

# export
ggsave(filename = "wind_plot.png", plot = wind_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(wind ~ time, data = daily_data)
summary_model <- summary(lm_model)
summary_model # p-value: < 2.2e-16, slope: -0.000243 per decade
```

## Max Storm Size: Combine Data and Plot

```{r}

# daily data
daily_data <- CanESM2_85_46779 %>% 
  # convert units back from kg/m^2/s to m then convert to inches
  mutate(precip_newunits = precip * (1000/86400) * 39.3701) %>% 
  # compute 3 day window of storm_size 
  arrange(time) %>%
  mutate(storm_size = rollapply(precip_newunits, width = 3, by = 3, FUN = sum, align = "right", fill = NA, partial = TRUE, na.rm = TRUE)) %>% 
  select(time, precip_newunits, storm_size) %>% 
  na.omit()

# trend line data
trend_data <- CanESM2_85_46779 %>% 
  # convert units back from kg/m^2/s to m then convert to inches
  mutate(precip_newunits = precip * (1000/86400) * 39.3701) %>% 
  # compute 3 day window of storm_size 
  arrange(time) %>%
  mutate(storm_size = rollapply(precip_newunits, width = 3, by = 3, FUN = sum, align = "right", fill = NA, partial = TRUE, na.rm = TRUE)) %>% 
  select(time, water_year, precip_newunits, storm_size) %>% 
  na.omit() %>% 
  # compute max storm size per year
  group_by(water_year) %>% 
  summarise(max_storm = max(storm_size)) %>% 
   mutate(time = as.Date(paste0(water_year, "-01-01")))
  
# plot

# Set Montserrat as the default font
theme_set(theme_bw(base_family = "Montserrat"))

precip_plot <- ggplot() +
  
  # max rh daily lines and fill inbetween
  geom_line(data = daily_data, aes(x = time, y = storm_size), color = "#EAE7E5") +
  
  # annotation for daily range
      # label
  annotate("text", x = as.Date("2065-01-01"), y = .05,
           label = "Storm Size Variation", color = '#606060', size = 5) +
      # line
  annotate("segment", 
           x = as.Date("2049-01-01"), xend = as.Date("2049-01-01"), 
           y = 0, yend = .17,
           color = "#606060",
           arrow = arrow(ends = "both", angle = 90, length = unit(.1, "cm"))) +
  
  # regression line
  geom_smooth(data = trend_data, aes(x = time, y = max_storm), color = '#CC4C34', size = 1, method = "lm", se = FALSE) +
  
  # labels for trend line
  annotate("text", x = as.Date("1969-01-01"), y = 0.065, 
           label = "Yearly Maximum Storm Size Trend", vjust = -1, color = '#CC4C34', size = 5) +
  
  #annotation for the trend line slope
 # annotate("text", x = as.Date("2000-01-01"), y = 0.1,
  #         label = "storm sizes projected to \n increase X inches* per decade", color = '#CC4C34', size = 4) +
  

  # theme
  theme_bw() + 
  labs(
    title = "\nSanta Barbara Yearly Maximum Rainstorm Size (inches every 3 days) Trends Upward",
    caption = "\nData Source: https://cal-adapt.org/data/download/ \n *p-value: 0.000326",
    subtitle = "Rainstorm Sizes Are Projected to Increase by 0.002 Inches* each Decade, on Average",
    y = "\nRainstorm Size (inches every 3 days)\n",
    x = "\nYear (1950 - 2100)"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 20, color = "#626262", hjust = 0),
    plot.subtitle = element_text(size = 15, color = "#626262", hjust = 0),
    axis.title.x = element_text(size = 13, color = "#626262"),
    axis.title.y = element_text(size = 13, color = "#626262"),
    axis.text = element_text(size = 11, color = "#626262"),
    plot.caption = element_text(size = 10, color = "#626262"),
    # remove vertical grid lines
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
  ) + 
  
  # add a dual axis
  scale_y_continuous(sec.axis = dup_axis())

# export
ggsave(filename = "precip_plot.png", plot = precip_plot, width = 13.3, height = 7.5, dpi = 300)

# significance testing
# for the yearly data since that's what we're plotting 
lm_model <- lm(max_storm ~ time, data = trend_data)
summary_model <- summary(lm_model)
summary_model
```

## Appendix: Other Trend Line Ideas

```{r}
# get a yearly average trend line from CanESM
# yearly_dat <- CanESM2_85_46779 %>% 
#   group_by(water_year) %>% 
#   summarise(max_temp = mean(max_temp), min_temp = mean(min_temp)) %>% 
#   mutate(water_year = as.Date(paste0(water_year, "-01-01"))) %>% 
#   mutate(max_temp = max_temp * 1.8 + 32) %>% 
#   mutate(min_temp = min_temp * 1.8 + 32)

# average for all 
# trend_data_t <- combined_df %>% 
#   group_by(water_year) %>% 
#   summarise(max_temp = mean(max_temp), min_temp = mean(min_temp)) %>% 
#   mutate(water_year = as.Date(paste0(water_year, "-01-01"))) %>% 
#   mutate(max_temp = max_temp * 1.8 + 32) %>% 
#   mutate(min_temp = min_temp * 1.8 + 32)

# trend_data <- combined_df %>% 
#   mutate(group = cut(time, breaks = "30 days")) %>% 
#   group_by(group) %>% 
#   summarise(max_temp = mean(max_temp), min_temp = mean(min_temp)) %>% 
#   mutate(max_temp = max_temp * 1.8 + 32) %>% 
#   mutate(min_temp = min_temp * 1.8 + 32) %>% 
#   mutate(time = as.Date(group)) %>% 
#   select(-group)
```

## 
